NURM: a tool to locate numerical differences in pipelines

### 2019-05-27 paper.tex

The remainder of this paper describes the NURM-tool in different parts 
including the way of capturing and representation of the provenance 
information, the method of clustering of different subject types, and 
process classification to characterize differences. We describe the HCP 
preprocessing pipelines as a popular neuroimaging project to test the 
NURM-tool and evaluate its functionality. Finally, we reported the 
experimental results which describe the processes that are responsible 
for the differences in the pipelines.

@@

Figure~\ref{fig:overview} shows an overview of our method.
For each tested condition, the
pipeline is containerized using Docker, and its interface is
described using Boutiques~\cite{glatard2017boutiques}. \reprozip
is added to the container images for provenance tracking. Our tool first
clusters the subjects based on the similarity of their process trees
produced by \reprozip (Fig.~\ref{fig:overview-cluster}). This results in a
reduced set of representative process trees to evaluate. The processes in
each representative process tree are labeled \tristan{Replace
``classification'' by ``labelling'' everyrwhere} from potential differences
found in their outputs produced in the tested conditions (Fig.~\ref{fig:overview-classify}).
To do that, a difference file is first produced that contains all the
files with differences across conditions
(Fig.~\ref{fig:overview-classify} - (A)\tristan{Add labels to the figure}).
Differences are identified simply by comparing file checksums. Using the
difference file, we then label processes in 4 categories (Fig.~\ref{fig:overview-classify} - (B)) depending
whether they create differences (represented in red), remove differences
(blue), neither create nor remove differences (green), or have uncertain
behavior (yellow). When processes with an uncertain behavior are found, the
pipeline in condition 2 is modified (as explained later) and re-executed. This
procedure iterates until all processes are labeled. 
\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=.7\columnwidth]{images/Slide1}
    \caption{Subject clustering.}
    \label{fig:overview-cluster}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=.75\columnwidth]{images/Slide2}
     \caption{Process classification.}
     \label{fig:overview-classify}
   \end{subfigure}
   \caption{Method overview}
   \label{fig:overview}
  \end{figure}

@@

NURM-tool classifies the processes into four 
categories based on the process tree and the \textit{difference} JSON file 
mentioned in the previous sections:
\begin{enumerate}
\item Processes that read and write files that do not have differences 
are \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have any differences but write files that have differences 
\emph{create} differences in the pipeline (Figure~\ref{fig:processes}.b).
\item Processes that read files 
that have differences and write files that do not have differences \emph{remove} 
differences from the pipeline (Figure~\ref{fig:processes}.c).
\item Processes that read files that have differences and write files that 
also have differences are 
\emph{unknown} (Figure~\ref{fig:processes}.d).
\end{enumerate}

@@

There are some issues that should be addressed to classify process 
properly, including capturing temporary files that have been removed 
during the execution, capturing files written by multiple processes, 
and unknown process.

\subsubsection{Capture temporary files} 
The classification of processes that read or 
write temporary files is uncertain when these files are deleted during 
the execution. To address this issue, we replace every process P that 
writes temporary files with a modification of P that first calls P and 
then backs up all its output files to a read-only directory. This 
replacement is done by modifying the Unix PATH variable to point to a 
directory containing the modified versions of the processes. This 
solution does not cover the temporary files that are removed by P 
itself; this is not a problem since these files do not play any role in 
the subsequent steps of the pipeline, by definition.

\subsubsection{Capture multiple write files} 
Files written by multiple processes also lead 
to unknown classification labels. To address this issue for a file F 
written by processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
(1) check that processes in \textbf{P} do not write concurrently to F, 
(2) we establish an order on \textbf{P} based on the creation timestamp 
of the processes, (3) we replace ever process $P_{i}$ in \textbf{P} by 
a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
directory. Thus, multiple versions of F are saved and used in the 
analysis.

@@

\subsubsection{Unknown process} 
Another challenge is to classify processes that 
read files that have differences and write files that also have differences. In 
such situations, it is not possible to determine from the pipeline 
results whether the process created differences, removed some differences, or was 
transparent. We label such processes \emph{unknown}.
To address this issue, we developed an iterative approach that 
consists of the following steps:

\begin{enumerate}
  \item Run the pipeline in conditions 1 and 2; classify the
    processes as \emph{transparent}, \emph{create differences},
    \emph{remove differences} or \emph{unknown}.
  \item If there are \emph{unknown} processes then: replace each 
  process P that create differences by a process Q that copies the results 
  produced by process P in condition 1 to the pipeline output in 
  condition 2. 
  \item Repeat steps 1 and 2 until there is no \emph{unknown} process left.
\end{enumerate}

The replacement of process P at step 2 is done by replacing P with a 
custom script produced by step 1. This custom script copies the results 
obtained in condition 1 if it is invoked with the arguments of a 
process that created differences, and calls P's original executable 
otherwise. The replacement is done through the PATH variable, as 
before. This algorithm converges to a process tree without any 
\emph{unknown} process after a finite number of iterations.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{images/iterative_modif}
  \caption{iteration of proposed modification.}
  \label{fig:iterations}
\end{figure}

Figure~\ref{fig:iterations} illustrates our iterative classification 
process for the example in Figure~\ref{fig:simple_script}. At every 
step, processes that created differences are shown in red, processes that 
removed differences are in blue, processes that propagated differences are in 
yellow, and other processes (transparent processes) are in green. As in 
Figure~\ref{fig:simple_script}, plain black edges represent the process 
tree and dashed edges represent file dependencies: green edges 
represent files with no differences, while red edges represent files with 
differences. Temporary files are not represented because they have been 
backed up as previously described.

The three steps in Figure~\ref{fig:iterations} correspond to the 
iterations of the classification algorithm. At step one, \texttt{2\#bet2} 
is classified as difference creator (red) as it produced files with differences 
from files without differences. \texttt{3\#fslstats} is classified as difference  
remover (blue) as it produced files without differences from files with 
differences. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
unknown (yellow) as they produce files with differences from files with 
differences.

@@

As a result of those 3 steps, the final process classification is: 
\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red), 
\texttt{3\#fslstats} is difference remover (blue) and the other processes 
are transparent.

@@

We executed the pipelines using Docker containers to simplify the 
deployment of different operating system versions on execution 
platforms. The Docker images were built for the HCP PreFreeSurfer 
pipeline v3.19.0 in 
CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for 
reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. 
We collected the provenance trace (tree of executed processes and files 
accessed in read or write mode) for each type of 
subjects(Table~\ref{table:data}) using system-call interception as 
provided by the \reprozip tool~\cite{Chirigati2016}.

@@

Figure \ref{fig:complete-graph} shows the annotated provenance graph 
of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
node in the graph represent an executed process in the pipeline. 
Processes that created errors are shown in red, processes that removed 
errors are in blue, and other processes are in green.  Squares denote 
processes for which the classification is uncertain, due to temporary 
files that were removed during the execution. Black edges link 
sub-processes to their parents while dashed edges denote file 
dependencies between processes (green edges: files with no errors; red 
edges: files with errors; yellow edges: temporary files).

>>>>

The remainder of this paper describes the NURM-tool in different parts 
including the way of capturing and representation of the provenance 
information, the method of clustering of different subject types, and 
process labelling to characterize differences. We describe the HCP 
preprocessing pipelines as a popular neuroimaging project to test the 
NURM-tool and evaluate its functionality. Finally, we reported the 
experimental results which describe the processes that are responsible 
for the differences in the pipelines.

@@

Figure~\ref{fig:overview} shows an overview of our method. For each 
tested condition, the pipeline is containerized using Docker, and its 
interface is described using Boutiques~\cite{glatard2017boutiques}. 
\reprozip is added to the container images for provenance tracking. Our 
tool first clusters the subjects based on the similarity of their 
process trees produced by \reprozip (Fig.~\ref{fig:overview-cluster}). 
This results in a reduced set of representative process trees to 
evaluate. The processes in each representative process tree are labeled 
from potential differences found in their outputs produced in the 
tested conditions (Fig.~\ref{fig:overview-classify}). To do that, a 
difference file is first produced that contains all the files with 
differences across conditions (Fig.~\ref{fig:overview-classify} - (A)). 
Differences are identified simply by comparing file checksums. 
Using difference file and process tree, we first capture temporary and 
overwritten files in condition 1. In addition, we set various 
breakpoints in the pipeline in condition 2 to check the output files 
produced by each process immediately after running. We then label processes in 4 
categories (Fig.~\ref{fig:overview-classify} - (B)) depending whether 
they create differences (represented in red), remove differences 
(blue), neither create nor remove differences (green), or have 
uncertain behavior (yellow). When processes that create differences are 
found, their written files in condition 2 are replaced (as explained 
later) and then re-labelling processes. This procedure 
continues until the end of the pipeline execution when all processes 
are labeled correctly.

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=.7\columnwidth]{images/Slide1}
    \caption{Subject clustering.}
    \label{fig:overview-cluster}
  \end{subfigure}
   \begin{subfigure}{\columnwidth}
    \centering
     \includegraphics[width=1\columnwidth]{images/Slide2}
     \caption{Process labelling.}
     \label{fig:overview-classify}
   \end{subfigure}
   \caption{Method overview}
   \label{fig:overview}
  \end{figure}

@@

NURM-tool labels the processes into four 
categories based on the process tree and the \textit{difference} JSON file 
mentioned in the previous sections:
\begin{enumerate}
\item Processes that read and write files that do not have differences 
are \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have any differences but write files that have differences 
\emph{create} differences in the pipeline (Figure~\ref{fig:processes}.b).
\item Processes that read files 
that have differences and write files that do not have differences \emph{remove} 
differences from the pipeline (Figure~\ref{fig:processes}.c).
\item Processes that read files that have differences and write files that 
also have differences are 
\emph{unknown} (Figure~\ref{fig:processes}.d).
\end{enumerate}

@@

There are some issues that should be addressed to label process 
properly, including capturing temporary files that have been removed 
during the execution, capturing files written by multiple processes, 
and unknown process.

\subsubsection{Capture temporary files} 
The labelling of processes that read or 
write temporary files is uncertain when these files are deleted during 
the execution. To address this issue, we replace every process P that 
writes temporary files with a modification of P that first calls P and 
then backs up all its output files to a read-only directory. This 
replacement is done by modifying the Unix PATH variable to point to a 
directory containing the modified versions of the processes. This 
solution does not cover the temporary files that are removed by P 
itself; this is not a problem since these files do not play any role in 
the subsequent steps of the pipeline, by definition.

\subsubsection{Capture multiple write files} 
Files written by multiple processes also lead 
to unknown labels. To address this issue for a file F 
written by processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
(1) check that processes in \textbf{P} do not write concurrently to F, 
(2) we establish an order on \textbf{P} based on the creation timestamp 
of the processes, (3) we replace ever process $P_{i}$ in \textbf{P} by 
a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
directory. Thus, multiple versions of F are saved and used in the 
analysis.

@@

we first capture temporary and overwritten files in condition 1
we set various 
breakpoints in the pipeline in condition 2 to check the output files 
produced by each process after running. We then label processes in 4 
categories. When processes that create differences are 
found, their written files in condition 2 are replaced (as explained 
later) and then re-label processes. This procedure (stop/resume) 
continues until the end of the pipeline execution when all processes 
are labeled correctly.

\subsubsection{Unknown process}

Another challenge is to label processes that read files that have 
differences and write files that also have differences. In such 
situations, it is not possible to determine from the pipeline results 
whether the process created differences, removed some differences, or 
was transparent. We label such processes \emph{unknown}.
To address this issue, we first set various breakpoints in the pipeline 
in condition 2. Breakpoints are set by replacing all the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent}, \emph{create differences}, \emph{remove 
differences} or \emph{unknown} immediately after running original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
condition 1 to the pipeline output in condition 2 if it is invoked with 
the arguments of a process that created differences. 
We developed an incremental approach that consists of the following steps:

\begin{enumerate}
  \item Run the pipeline in conditions 2; after running each process 
  label it as \emph{transparent}, \emph{create differences} or
  \emph{remove differences}.

  \item If there is \emph{create differences} processes then: 
  copies the results produced by the same process in condition 1 to the 
  pipeline output in condition 2.

  \item Continue steps 1 and 2 until the end of the pipeline execution.
\end{enumerate}

This algorithm converges to a process tree without any 
\emph{unknown} process at the end of the pipeline execution.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{images/iterative_modif}
  \caption{A simple example of steps of the proposed approach.}
  \label{fig:iterations}
\end{figure}

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. At every 
step, processes that created differences are shown in red, processes that 
removed differences are in blue, processes that propagated differences are in 
yellow, and other processes (transparent processes) are in green. As in 
Figure~\ref{fig:simple_script}, plain black edges represent the process 
tree and dashed edges represent file dependencies: green edges 
represent files with no differences, while red edges represent files with 
differences. Temporary files are not represented because they have been 
backed up as previously described.

The three steps in Figure~\ref{fig:iterations} correspond to the 
steps of the labelling algorithm. At step one, \texttt{2\#bet2} 
is classified as difference creator (red) as it produced files with differences 
from files without differences. \texttt{3\#fslstats} is classified as difference  
remover (blue) as it produced files without differences from files with 
differences. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
unknown (yellow) as they produce files with differences from files with 
differences.

@@

As a result of those 3 steps, the final process labelling is: 
\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red), 
\texttt{3\#fslstats} is difference remover (blue) and the other processes 
are transparent.

@@

We executed the pipelines using Docker containers to simplify the 
deployment of different operating system versions on execution 
platforms. The Docker images were built for the HCP PreFreeSurfer 
pipeline v3.19.0 in 
CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for 
reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}.

@@

Figure \ref{fig:complete-graph} shows the annotated provenance graph 
of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
node in the graph represent an executed process in the pipeline. 
Processes that created errors are shown in red, processes that removed 
errors are in blue, and other processes are in green.  Squares denote 
processes for which the labelling is uncertain, due to temporary 
files that were removed during the execution. Black edges link 
sub-processes to their parents while dashed edges denote file 
dependencies between processes (green edges: files with no errors; red 
edges: files with errors; yellow edges: temporary files).
